---
title: "Crosstalk Example: GP Practice Data"
output: html_document
---

```{r setup, include=FALSE}
# Load necessary libraries
# Ensure these are installed: `install.packages(c("crosstalk", "plotly", "DT", "dplyr", "tidyr"))`
library(crosstalk)
library(plotly)
library(DT)
library(dplyr)
library(tidyr) # Only used for generating the dummy data

# Set seed for reproducibility
set.seed(123)
```

## Introduction

This document demonstrates how to use the `crosstalk` package in R Markdown to create linked interactive components. We will generate dummy data for three GP practices, showing a monthly value over a year. A filter control will allow selecting specific practices, and both the line plot and the data table below will update accordingly.

## Data Generation

First, we generate some sample data representing monthly values for three GP practices.

```{r generate_data}
# Define GP practices
practices <- c("Practice A", "Practice B", "Practice C")

# Create a sequence of months for one year
months <- seq(as.Date("2023-01-01"), as.Date("2023-12-01"), by = "month")

# Create a data frame combining practices and months
gp_data <- expand_grid(practice_name = practices, month = months)

# Add a random value for each practice/month combination
# Simulate some variation and trends
gp_data <- gp_data |>
  arrange(practice_name, month) |>
  group_by(practice_name) |>
  mutate(
    # Add some baseline variation and a slight trend component
    value = round(50 + cumsum(rnorm(n(), mean = 0, sd = 5)) + seq(0, 5, length.out = n()), 1),
    # Ensure value is not negative
    value = pmax(0, value)
  ) |>
  ungroup() |>
  # Format month for better display
  mutate(month_display = format(month, "%b-%y"))
```

## Interactive Visualization and Table

Now, we wrap the data in `SharedData` from the `crosstalk` package. This allows different widgets and plots to react to selections made on the data.

```{r crosstalk_setup}
# Wrap the data frame in SharedData
# Use practice_name as the key for linking
shared_gp_data <- SharedData$new(gp_data, key = ~practice_name, group = "gp_selection")
```

We can now create the interactive components: a filter for the GP practice, a Plotly line plot, and a DT datatable.

```{r interactive_filter}
# Create the filter control for GP practice
# Allows selecting one or more practices
filter_select(
  id = "practice_filter",
  label = "Select GP Practice(s):",
  sharedData = shared_gp_data,
  group = ~practice_name # Filter based on the practice_name column
)
```

```{r interactive_plotly}
# Create the Plotly line plot
# It uses the shared data, so it will update based on the filter
plot_gp <- plot_ly(
  shared_gp_data,
  x = ~month,
  y = ~value,
  color = ~practice_name,
  type = "scatter",
  mode = "lines+markers",
  hoverinfo = "text",
  text = ~ paste("Practice:", practice_name, "<br>Month:", month_display, "<br>Value:", value)
) |>
  layout(
    title = "Monthly Value per GP Practice",
    xaxis = list(title = "Month"),
    yaxis = list(title = "Value")
  )
```

```{r interactive_datatable}
# Create the DT datatable
# It also uses the shared data and will update
table_gp <- datatable(
  shared_gp_data,
  rownames = FALSE, # Don't show row numbers
  colnames = c("Practice Name", "Month", "Value", "Month Display"), # Nicer column names
  filter = "none", # Disable individual column filters as we have the main filter
  options = list(
    pageLength = 10, # Show 10 rows per page
    autoWidth = TRUE
  )
)
```

```{r display_elements}
# Display the plot and table
# They will react to the filter selection above
bscols(
  plot_gp, 
  table_gp, 
  widths = c(6, 6), # Total 'columns' is always 12 
  device = "lg" # Use columns on a 'large' screen but collapse on a smaller one.
  )  
```

## Explanation

1.  **Data Preparation**: We created a tidy data frame (`gp_data`) with columns for practice name, month, and a simulated value.
2.  **Shared Data**: The `SharedData$new()` function wraps our data frame. We specify `key = ~practice_name` so that `crosstalk` knows which variable links selections across different components (the filter, plot, and table). The `group` argument ensures that widgets intended to work together share the same selection scope.
3.  **Filter Control**: `filter_select()` creates a drop-down/select input. It's linked to the `shared_gp_data` via the `sharedData` argument and targets the `practice_name` column via the `group` argument.
4.  **Plotly Plot**: `plot_ly()` is called directly on the `shared_gp_data` object. Plotly automatically understands how to interact with `SharedData` objects, filtering the data based on the current selection from the `filter_select` widget.
5.  **DT Table**: Similarly, `datatable()` is called on the `shared_gp_data` object. It filters its display based on the selection.
6.  **Layout**: `bscols()` (from `crosstalk`) arranges the plot and table side-by-side for a cleaner layout.

When you select one or more practices from the filter drop-down, both the plot and the table will instantly update to show only the data for the selected practices.
